Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Final State with Multivariate Techniques

### 2015-12-14 paper/sec-mcsamples.tex
implementing Daniela's comments

Higgs pair production is simulated at leading order (LO) using
{\tt MadGraph5\_aMC@NLO}~\cite{Alwall:2014hca}.
%
We use a tailored  model~\cite{Maltoni:2014eza}
for gluon-fusion Higgs boson pair production
which includes mass effects
from the
exact form factors for the top-quark triangle and box
loops~\cite{Plehn:1996wb}.
%
Equivalent results can be obtained using
the recently available functionalities
for the calculation of loop-induced processes
in {\tt MadGraph5\_aMC@NLO}~\cite{Hirschi:2015iia}.
%
The calculation is performed in the
$n_f$=4 scheme,  accounting for  $b$-quarks mass effects,
and the renormalization and factorization
scales are taken to be $\mu_F=\mu_R=H_T/2$.
%
For the input parton distribution functions (PDFs), we 
adopt the NNPDF 3.0 $n_f = 4$ LO set~\cite{Ball:2014uwa} with
$\alpha_s(m_Z^2)=0.118$,
interfaced via {\tt LHAPDF6}~\cite{Buckley:2014ana}.
%
The Higgs boson couplings
and branching ratios are set to their SM values,
and its mass is taken to be
$m_h=125$ GeV~\cite{Aad:2014aba,Khachatryan:2014jba,Aad:2015zhl}.
%
In the SM, the Higgs trilinear coupling
is given by $\lambda=m_h^2/2v^2$, with $m_h=125$ GeV
the Higgs boson mass and $v\simeq 246$ GeV the Higgs vacuum expectation
value (vev).
%

@@

In Fig.~\ref{fig:hhFeyn} we show representative Feynman diagrams
    for LO Higgs pair production in gluon fusion.
    %
    The non-trivial interplay between the diagram with a heavy quark box
    and that of the triangle, that can lead to either
    constructive or destructive interference,
    complicates the extraction of
    the trilinear coupling
    $\lambda$ from the measurement of the Higgs pair
    production cross-section.
    %
    Higher order corrections~\cite{deFlorian:2013jea,Frederix:2014hta}
    are dominated by gluon radiation
    from either the initial state gluons or from the heavy quark loops.

@@

The LO cross-sections for
the background samples have been rescaled so that the integrated
distributions reproduce known higher-order QCD results.
%
For the $4b$ and $2b2j$ samples, a NLO/LO $K$-factor has been determined
using {\tt MadGraph5\_aMC@NLO}~\cite{Alwall:2014hca}, which turns out to be 1.6 and 1.3
respectively.
%
For the $4j$ sample, we rescale the LO
using the {\tt BLACKHAT}~\cite{Bern:2011ep}
calculation, that indicate
a NLO/LO $K$-factor of 0.6.
%
Finally, the LO cross-section for $t\bar{t}$ production has been rescaled
to match the NNLO+NNLL calculation of Ref.~\cite{Czakon:2013goa}, which leads
to a $K$-factor of 1.4.
%
The $K$-factors that we use to rescale
the signal background samples are collected in
Table~\ref{tab:samples}.

@@

While it is beyond the scope of this work to perform a full
detector simulation, it is important to include an estimate of detector
effects in the analysis, in particular for the finite energy
and angular resolutions, which directly
degrade the reconstruction of important kinematical variables, such as
the invariant mass of the Higgs candidates.
%
In this work, we simulate the finite energy resolution of the ATLAS and CMS
hadronic calorimeters by applying a Gaussian smearing of the transverse
momentum $p_T$ of all
final-state particles, before jet clustering,
with mean zero and standard deviation $\sigma_E$, that is,
%
\be
\label{eq:smearing}
p_T^{(i)} \, \to \, p_T^{(i)\prime}= \lp 1+ r_i\cdot\sigma_E \rp\, p_T^{(i)} \, , \quad
i=1,\ldots,N_{\rm part} \, ,
\ee
with $r_i$ an univariate Gaussian random number, different for each
of the $N_{\rm part}$ particles in the event.
%
We take as baseline value for the transverse momentum smearing a
factor of $\sigma_E=5\%$.
%

>>>>

Higgs pair production is simulated at leading order (LO) using
{\tt MadGraph5\_aMC@NLO}~\cite{Alwall:2014hca}.
%
We use a tailored  model~\cite{Maltoni:2014eza}
for gluon-fusion Higgs boson pair production
which includes mass effects
from the
exact form factors for the top-quark triangle and box
loops~\cite{Plehn:1996wb}.
%
Equivalent results can be obtained using
the recently available functionalities
for the calculation of loop-induced processes
in {\tt MadGraph5\_aMC@NLO}~\cite{Hirschi:2015iia}.
%
The calculation is performed in the
$n_f$=4 scheme,  accounting for  $b$-quarks mass effects,
and the renormalization and factorization
scales are taken to be $\mu_F=\mu_R=H_T/2$,
with
\be
H_T\equiv \sum_i \sqrt{p_{T,i}^2+m_i^2} \, ,
\ee
the scalar sum of the
transverse masses of all final state particles.
%
For the input parton distribution functions (PDFs), we 
adopt the NNPDF 3.0 $n_f = 4$ LO set~\cite{Ball:2014uwa} with
$\alpha_s(m_Z^2)=0.118$,
interfaced via {\tt LHAPDF6}~\cite{Buckley:2014ana}.
%
The Higgs boson couplings
and branching ratios are set to their SM values,
and its mass is taken to be
$m_h=125$ GeV~\cite{Aad:2014aba,Khachatryan:2014jba,Aad:2015zhl}.
%
In the SM, the Higgs trilinear coupling
is given by $\lambda=m_h^2/2v^2$, with $m_h=125$ GeV
the Higgs boson mass and $v\simeq 246$ GeV the Higgs vacuum expectation
value (vev).
%

@@

In Fig.~\ref{fig:hhFeyn} we show representative Feynman diagrams
    for LO Higgs pair production in gluon fusion.
    %
    The non-trivial interplay between the diagram with a heavy quark box
    and that of the triangle loop,
    that can lead to either
    constructive or destructive interference,
    complicates the extraction of
    the trilinear coupling
    $\lambda$ from the measurement of the Higgs pair
    production cross-section.
    %
    Higher order corrections~\cite{deFlorian:2013jea,Frederix:2014hta}
    are dominated by gluon radiation
    from either the initial state gluons or from the heavy quark loops.

@@

The LO cross-sections for
the background samples have been rescaled so that the integrated
distributions reproduce known higher-order QCD results.
%
For the $4b$ and $2b2j$ samples, a NLO/LO $K$-factor has been determined
using {\tt MadGraph5\_aMC@NLO}~\cite{Alwall:2014hca}, which turns out to be 1.6 and 1.3
respectively.
%
For the $4j$ sample, we rescale the LO
using the {\tt BLACKHAT}~\cite{Bern:2011ep}
calculation, that indicates
a NLO/LO $K$-factor of 0.6.
%
Finally, the LO cross-section for $t\bar{t}$ production has been rescaled
to match the NNLO+NNLL calculation of Ref.~\cite{Czakon:2013goa}, which leads
to a $K$-factor of 1.4.
%
The $K$-factors that we use to rescale
the signal background samples are collected in
Table~\ref{tab:samples}.

@@

While it is beyond the scope of this work to perform a full
detector simulation, it is important to include an estimate of detector
effects in the analysis, in particular for the finite energy
and angular resolutions, which directly
degrade the reconstruction of important kinematical variables, such as
the invariant mass of the Higgs candidates.
%
In this work, we simulate the finite energy resolution of the ATLAS and CMS
hadronic calorimeters by applying a Gaussian smearing of the transverse
momentum $p_T$ for all
final-state particles, before jet clustering,
with mean zero and standard deviation $\sigma_E$, that is,
%
\be
\label{eq:smearing}
p_T^{(i)} \, \to \, p_T^{(i)\prime}= \lp 1+ r_i\cdot\sigma_E \rp\, p_T^{(i)} \, , \quad
i=1,\ldots,N_{\rm part} \, ,
\ee
with $r_i$ an univariate Gaussian random number, different for each
of the $N_{\rm part}$ particles in the event.
%
We take as baseline value for the transverse momentum smearing a
factor of $\sigma_E=5\%$.
%
